### 2024-04-21 paper/paper.qmd
updated the data section of paper

---
title: "Formula 1 2022 Regulation Changes"
subtitle: "How did the 2022 F1 Regulations change the sport for better and for worse"
author: 
  - Michael Fang
thanks: "Code and data are available at: https://github.com/fanger2791/F1_Regulations."
date: today
date-format: long
abstract: "As Formula 1's new era of hybrid cars develops, a fresh range of rules and regulations have been put in place for the teams to follow attempting to create a closer, more competitive championship. This paper will investigate the impact of the 2022 Formula 1 regulations, designed to intensify the competitive landscape and elevate the racing experience through technical and fiscal adjustments. It examines lap times and overtaking dynamics, while the analysis zeroes in on the performance trajectories of the sport’s leading contenders: Max Verstappen, Lewis Hamilton, and Charles Leclerc across varied circuit types from 2021 to 2022. It focuses on whether these regulatory modifications have indeed leveled the playing field, creating tighter competition and mitigating the performance differences between teams or have increased the gap even more. This research not only examines the immediate effects of the 2022 changes but also considers the balance between innovation and competition in Formula 1, thereby contributing to the ongoing dialogue on how to shape the future of this high-stakes sport."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

# Introduction

Formula 1 (F1) is the pinnacle of motorsport, recognized worldwide for its high-speed and high-tech racing that pushes the limits of automotive engineering and driver skill to the limit. Established in 1950, it has evolved into a global phenomenon, comprising of a series of races, known as Grands Prix, held on different tracks and street circuits across various continents. The sport features single-seat, open-wheel cars powered by cutting-edge hybrid technology, emphasizing aerodynamic performance, energy recovery systems, and fuel efficiency. Teams, both factory-backed and independent, compete in a Constructors' Championship, while drivers compete for the Drivers' Championship, making F1 not just a test of technical innovation but also human endurance and strategy.

In 2022, Formula 1 underwent one of its most significant regulatory overhauls in the history of the sport, a change motivated by the desire to enhance competitive balance and improve racing quality. This transformation came at a critical juncture, as the world of Formula 1 sought to address growing concerns over the technological disparities among teams, and the often predictable nature of race outcomes. The regulations introduced for the 2022 season aimed to tackle these challenges by mandating a new car design philosophy, overhauling the aerodynamics to facilitate closer racing, and implementing cost caps to level the playing field among the competing teams.

The core of the 2022 regulations centered around a shift to ground-effect aerodynamics, a move designed to reduce the reliance on turbulent airflows generated by traditional wing-based designs and thereby minimize the performance degradation experienced by cars when following closely behind competitors. Additionally, the introduction of a budget cap aimed to reduce the escalating arms race in spending that had seen the wealthiest teams dominate the sport. These changes were complemented by adjustments to race weekend formats, marking a comprehensive approach to reforming the sport.

However, the transition was not without its challenges and controversies. Critics and experts argued that while the intention was to even the competition, the rapid and radical nature of these changes risked worsening existing disparities, at least in the short term, as teams adapted at different rates to the new regulations. Furthermore, concerns were raised about the potential for unintended consequences, such as increased racing incidents due to closer pack racing, and the possibility that the budget cap could stifle technological innovation.

Despite these debates, preliminary findings suggest that the 2022 regulations have had a profound impact on Formula 1. Races have generally become more competitive, with a wider array of teams and drivers finding themselves in contention for podium finishes. The budget cap has also initiated a more strategic allocation of resources among teams, potentially leveling the competitive field over the longer term. The paper's estimand is whether the 2022 Formula 1 regulations led to a significant increase in lap time in comparison to the previous time among the top drivers across different types of circuits but also increased the number of successful overtakes during races. We aim to quantify the impact of the 2022 Formula 1 regulatory changes, providing a clear basis for evaluating their success in enhancing the sport's competitive balance and the quality of racing.

The remainder of the paper is organized as follows: @sec-data details the methodology and source data, including the use of the f1dataR package. @sec-results provides an analysis of the data with an emphasis on changes in the laptime and number of overtakes completed between 2021 and 2022 F1 cars. @sec-discussion discusses the specific challenges that came with the new regulations and on the balance between the intended objectives and the actual outcomes, offering recommendations for future regulatory adjustments.

# Data {#sec-data}

## Source Data

The f1dataR [@f1dataR] package in R is designed specifically for Formula 1 enthusiasts and data analysts who wish to examine the statistical and historical aspects of Formula 1 racing. This package provides a comprehensive setf functions that facilitate the extraction and manipulation of Formula 1 data directly from the Ergast API, which is a widely recognized source for F1 statistics. With f1dataR, users can easily access detailed race results, driver standings, constructor details, and much more, across the history of F1 competitions.

## 2021 Standings vs 2022 Standings

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-standings2022
#| tbl-cap: F1 Driver Standings 2022
library(tidyverse)
library(readr)
library(dplyr)
library(kableExtra)

standings_2022 <- read_csv("../data/analysis_data/standings_2022_cleaned.csv")

standings_2022 |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-standings2021
#| tbl-cap: F1 Driver Standings 2021
library(tidyverse)
library(readr)
library(dplyr)
library(kableExtra)

standings_2021 <- read_csv("../data/analysis_data/standings_2021_cleaned.csv")

standings_2021 |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

The Formula 1 driver standings from 2021 to 2022 as seen in @tbl-standings2022 and @tbl-standings2021 show a notable progression and shift in team and driver performance. In 2021, Max Verstappen secured the championship with a 7.5-point lead over Lewis Hamilton, with 10 wins to Hamilton's 8. Mercedes appeared strong, with both Hamilton and Bottas in the top three. Come 2022, Verstappen further solidified his dominance by extending his lead to a substantial 146 points and achieving a record 15 wins in a single season. Notably, Ferrari improved their standing with Leclerc finishing second, and Red Bull maintained their position as the team to beat with Perez also featuring in the top three. The 2022 season's F1 regulations, which introduced major changes to the car design with an emphasis on ground-effect aerodynamics, appear to have shaken up the order slightly. This aimed to allow cars to follow each other more closely and to increase overtaking opportunities. The impact of these changes can be inferred from the reshuffling seen in team performance. Mercedes, for example, slipped from the top spot as Ferrari emerged as a stronger challenger to Red Bull. The midfield battle also saw shifts, with drivers like Ocon and Alonso gaining points, indicating that the new regulations might have provided an opportunity for teams to reset and challenge the established order.

## Method of Analysis

To analyze the impact of track characteristics on Formula 1 car performance, specifically comparing the lap times and speeds at the Saudi Arabia Jeddah Corniche Circuit (a top-speed track) and the Netherlands' Circuit Zandvoort (a high-downforce track) as seen in @tbl-circuits for the 2021 and 2022 seasons, a structured approach will be employed. Initially, data will be gathered using the f1dataR package, which extracts detailed information about lap times and speeds from the Ergast API. This dataset will include all drivers' performances during both qualifying and race sessions.

The analysis will begin with descriptive statistics to outline the basic trends in the ranges of lap times and top speeds for each session across both tracks and years. This provides a preliminary view of the data and helps identify any outliers or anomalies. For deeper insights, a comparative analysis will be conducted between the two tracks and between the two years. This involves comparing average speeds and lap times to determine how the distinct characteristics of each track influence performance and how performance metrics have evolved from one year to the next.

Finally, the results will be visually represented through various graphs and plots to illustrate variability and distribution. This comprehensive approach ensures that the analysis not only provides quantitative insights into F1 car performance under varying track conditions but also presents these findings in an accessible format, aiding in the interpretation and discussion of how different track characteristics influence race strategy and outcomes.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-circuits
#| tbl-cap: F1 Circuits Analysis
library(tidyverse)
library(kableExtra)

variable_data <- data.frame( 
  Circuit_Name = c("Saudi Arabia", 
               "Netherlands"), 
  Track_Type = c("Top Speed",
                  "High Downforce"), 
  Tyre_Wear = c("High", 
              "Low")
)
variable_data |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

## Data Limitations

While the f1dataR [@f1dataR] package offers valuable resources for analyzing Formula 1 data, it does come with certain limitations that users need to consider. First, the quality and completeness of the data are dependent on the Ergast API, which means any missing or inaccurately recorded data in the API will be reflected in the outputs obtained through f1dataR. Additionally, the API's update frequency might not align with real-time event conclusions, potentially leading to delays in accessing the most current race data. The historical data provided may also have gaps, particularly from the earlier years of Formula 1 racing, which could limit analyses of long-term trends or comparisons over extended periods. Moreover, the package may not include certain types of data, such as detailed telemetry or in-depth weather conditions, which could be crucial for more nuanced analyses or predictive modeling. Therefore, while f1dataR facilitates access to a wealth of F1 data, users must be mindful of these constraints when conducting comprehensive or real-time analyses.

## Data Cleaning

R [@citeR] was the language and environment used for this paper as well as throughout the data cleaning process, with different packages such as tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], tidyr [@tidyr], knitr [@knitr], janitor [@janitor], lubridate [@lubridate] packages.

The data cleaning process for these datasets involves several steps to ensure the data is accurate and suitable for analysis. Initially, the dataset's structure is reviewed to verify data types and identify any inconsistencies. This is crucial for understanding the data's composition and guiding subsequent transformations. Key transformations include converting relevant columns to appropriate data types, such as changing lap times to numeric values for analytical operations and categorizing seasons as factors for grouped analyses. Such transformations are essential for maintaining data integrity and ensuring computational accuracy. A critical cleaning step involves filtering out anomalies, specifically laps exceeding 100 seconds, based on the assumption that these represent non-standard race conditions like pit stops or errors. This step helps focus the dataset on typical racing performance. Further, the dataset undergoes a purging of incomplete records, removing any rows with missing data to enhance the robustness of the statistical analysis. Finally, renaming columns to more descriptive terms facilitates easier understanding and manipulation of the data, aligning with best practices in data documentation and usability.

# Results {#sec-results}

```{r}
#| label: fig-maxjeddah
#| fig-cap: Max Verstappen Lap Times in Jeddah - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ver_jeddah_2022 <- read_csv("../data/analysis_data/verstappen_laps_jeddah_2022_cleaned.csv")
ver_jeddah_2021 <- read_csv("../data/analysis_data/verstappen_laps_jeddah_2021_cleaned.csv")

# Add a season column to each dataframe
ver_jeddah_2021$Season <- '2021'
ver_jeddah_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ver_jeddah_2021, ver_jeddah_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "blue", "2022" = "red"))

```

```{r}
#| label: fig-lewisjeddah
#| fig-cap: Lewis Hamilton Lap Times in Jeddah - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ham_jeddah_2022 <- read_csv("../data/analysis_data/hamilton_laps_jeddah_2022_cleaned.csv")
ham_jeddah_2021 <- read_csv("../data/analysis_data/hamilton_laps_jeddah_2021_cleaned.csv")

# Add a season column to each dataframe
ham_jeddah_2021$Season <- '2021'
ham_jeddah_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ham_jeddah_2021, ham_jeddah_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))
```

```{r}
#| label: fig-lecjeddah
#| fig-cap: Charles Leclerc Lap Times in Jeddah - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

lec_jeddah_2022 <- read_csv("../data/analysis_data/leclerc_laps_jeddah_2022_cleaned.csv")
lec_jeddah_2021 <- read_csv("../data/analysis_data/leclerc_laps_jeddah_2021_cleaned.csv")

# Add a season column to each dataframe
lec_jeddah_2021$Season <- '2021'
lec_jeddah_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(lec_jeddah_2021, lec_jeddah_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))
```

```{r}
#| label: fig-maxdutch
#| fig-cap: Max Verstappen Lap Times in the Netherlands - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ver_dutch_2022 <- read_csv("../data/analysis_data/verstappen_laps_dutch_2022_cleaned.csv")
ver_dutch_2021 <- read_csv("../data/analysis_data/verstappen_laps_dutch_2021_cleaned.csv")

# Add a season column to each dataframe
ver_dutch_2021$Season <- '2021'
ver_dutch_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ver_dutch_2021, ver_dutch_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "blue", "2022" = "red"))

```

```{r}
#| label: fig-lewisdutch
#| fig-cap: Lewis Hamilton Lap Times in the Netherlands - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ham_dutch_2022 <- read_csv("../data/analysis_data/hamilton_laps_dutch_2022_cleaned.csv")
ham_dutch_2021 <- read_csv("../data/analysis_data/hamilton_laps_dutch_2021_cleaned.csv")

# Add a season column to each dataframe
ham_dutch_2021$Season <- '2021'
ham_dutch_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ham_dutch_2021, ham_dutch_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))
```

```{r}
#| label: fig-lecdutch
#| fig-cap: Charles Leclerc Lap Times in the Netherlands - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

lec_dutch_2022 <- read_csv("../data/analysis_data/leclerc_laps_dutch_2022_cleaned.csv")
lec_dutch_2021 <- read_csv("../data/analysis_data/leclerc_laps_dutch_2021_cleaned.csv")

# Add a season column to each dataframe
lec_dutch_2021$Season <- '2021'
lec_dutch_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(lec_dutch_2021, lec_dutch_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))
```

```{r}
#| label: fig-overtakes
#| fig-cap: Number of Overtakes Seen on TV 2021 v 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

data <- read.csv("../data/raw_data/Overtaking_Season_Averages.csv")

# Filter data for the years 2021 and 2022
filtered_data <- data[data$Season %in% c(2021, 2022), ]

ggplot(filtered_data, aes(x = as.factor(Season), y = Overtakes.on.TV, fill = "Overtakes on TV")) +
    geom_bar(stat = "identity", color = "black", fill = "red") +
    labs(title = "Overtakes on TV (2021-2022)",
         x = "Season", y = "Number of Overtakes on TV") +
    theme_minimal()
```

```{r}
#| label: fig-verspeed-jeddah
#| fig-cap: Verstappen Speed Comparison - Qualifying Jeddah 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ver_telemetry_data_jeddah_2022 <- read.csv("../data/raw_data/verstappen_telemetry_jeddah.csv")
ver_telemetry_data_jeddah_2021 <- read.csv("../data/raw_data/verstappen_telemetry_jeddah_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
ver_telemetry_data_jeddah_2022$season <- "2022"
ver_telemetry_data_jeddah_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(ver_telemetry_data_jeddah_2022, ver_telemetry_data_jeddah_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "blue"))

```

```{r}
#| label: fig-hamspeed-jeddah
#| fig-cap: Hamilton Speed Comparison - Qualifying Jeddah 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ham_telemetry_data_jeddah_2022 <- read.csv("../data/raw_data/hamilton_telemetry_jeddah.csv")
ham_telemetry_data_jeddah_2021 <- read.csv("../data/raw_data/hamilton_telemetry_jeddah_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
ham_telemetry_data_jeddah_2022$season <- "2022"
ham_telemetry_data_jeddah_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(ham_telemetry_data_jeddah_2022, ham_telemetry_data_jeddah_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))
```
```{r}
#| label: fig-lecspeed-jeddah
#| fig-cap: Leclerc Speed Comparison - Qualifying Jeddah 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

lec_telemetry_data_jeddah_2022 <- read.csv("../data/raw_data/leclerc_telemetry_jeddah.csv")
lec_telemetry_data_jeddah_2021 <- read.csv("../data/raw_data/leclerc_telemetry_jeddah_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
lec_telemetry_data_jeddah_2022$season <- "2022"
lec_telemetry_data_jeddah_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(lec_telemetry_data_jeddah_2022, lec_telemetry_data_jeddah_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))
```

# Discussion {#sec-discussion}

\newpage

# References