### 2024-04-23 paper/paper.qmd
added conclusion and more references

---
title: "Formula 1 2022 Regulation Changes"
subtitle: "How did the 2022 F1 Regulations change the sport"
author: 
  - Michael Fang
thanks: "Code and data are available at: https://github.com/fanger2791/F1_Regulations."
date: today
date-format: long
abstract: "As Formula 1's new era of hybrid cars develops, a fresh range of rules and regulations have been put in place for the teams to follow attempting to create a closer, more competitive championship. This paper will investigate the impact of the 2022 Formula 1 regulations, designed to intensify the competitive landscape and elevate the racing experience through technical and fiscal adjustments. It examines lap times and overtaking dynamics, while the analysis zeroes in on the performance trajectories of the sport’s leading contenders: Max Verstappen, Lewis Hamilton, and Charles Leclerc across varied circuit types from 2021 to 2022. It focuses on whether these regulatory modifications have indeed leveled the playing field, creating tighter competition and mitigating the performance differences between teams or have increased the gap even more. This research not only examines the immediate effects of the 2022 changes but also considers the balance between innovation and competition in Formula 1, thereby contributing to the ongoing dialogue on how to shape the future of this high-stakes sport."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

# Introduction

Formula 1 (F1) is the pinnacle of motorsport, recognized worldwide for its high-speed and high-tech racing that pushes the limits of automotive engineering and driver skill to the limit. Established in 1950, it has evolved into a global phenomenon, comprising of a series of races, known as Grands Prix, held on different tracks and street circuits across various continents. The sport features single-seat, open-wheel cars powered by cutting-edge hybrid technology, emphasizing aerodynamic performance, energy recovery systems, and fuel efficiency. Teams, both factory-backed and independent, compete in a Constructors' Championship, while drivers compete for the Drivers' Championship, making F1 not just a test of technical innovation but also human endurance and strategy.

In 2022, Formula 1 underwent one of its most significant regulatory overhauls in the history of the sport, a change motivated by the desire to enhance competitive balance and improve racing quality. This transformation came at a critical juncture, as the world of Formula 1 sought to address growing concerns over the technological disparities among teams, and the often predictable nature of race outcomes. The regulations introduced for the 2022 season aimed to tackle these challenges by mandating a new car design philosophy, overhauling the aerodynamics to facilitate closer racing, and implementing cost caps to level the playing field among the competing teams.

The core of the 2022 regulations centered around a shift to ground-effect aerodynamics, a move designed to reduce the reliance on turbulent airflows generated by traditional wing-based designs and thereby minimize the performance degradation experienced by cars when following closely behind competitors. Additionally, the introduction of a budget cap aimed to reduce the escalating arms race in spending that had seen the wealthiest teams dominate the sport. These changes were complemented by adjustments to race weekend formats, marking a comprehensive approach to reforming the sport.

However, the transition was not without its challenges and controversies. Critics and experts argued that while the intention was to even the competition, the rapid and radical nature of these changes risked worsening existing disparities, at least in the short term, as teams adapted at different rates to the new regulations. Furthermore, concerns were raised about the potential for unintended consequences, such as increased racing incidents due to closer pack racing, and the possibility that the budget cap could stifle technological innovation.

Despite these debates, preliminary findings suggest that the 2022 regulations have had a profound impact on Formula 1. Races have generally become more competitive, with a wider array of teams and drivers finding themselves in contention for podium finishes. The budget cap has also initiated a more strategic allocation of resources among teams, potentially leveling the competitive field over the longer term. The paper's estimand is whether the 2022 Formula 1 regulations led to a significant increase in lap time and overall top speed in comparison to the previous time among the top drivers across different types of circuits but also increased the number of successful overtakes during races. We aim to quantify the impact of the 2022 Formula 1 regulatory changes, providing a clear basis for evaluating their success in enhancing the sport's competitive balance and the quality of racing.

The remainder of the paper is organized as follows: @sec-data details the methodology and source data, including the use of the f1dataR package. @sec-results provides an analysis of the data with an emphasis on changes in the laptime, speed and number of overtakes completed between 2021 and 2022 F1 cars. @sec-discussion discusses the specific challenges that came with the new regulations and on the balance between the intended objectives and the actual outcomes, offering recommendations for future regulatory adjustments.

# Data {#sec-data}

## Source Data

The f1dataR [@f1dataR] package in R is designed specifically for Formula 1 enthusiasts and data analysts who wish to examine the statistical and historical aspects of Formula 1 racing. This package provides a comprehensive set of functions that facilitate the extraction and manipulation of Formula 1 data directly from the Ergast API, which is a widely recognized source for F1 statistics. With f1dataR, users can easily access detailed race results, driver standings, constructor details, and much more, across the history of F1 competitions.

## 2021 Standings vs 2022 Standings

The Formula 1 driver standings from 2021 to 2022 as seen in @tbl-standings2022 and @tbl-standings2021 show a notable progression and shift in team and driver performance. In 2021, Max Verstappen secured the championship with a 7.5-point lead over Lewis Hamilton, with 10 wins to Hamilton's 8. Mercedes appeared strong, with both Hamilton and Bottas in the top three. Come 2022, Verstappen further solidified his dominance by extending his lead to a substantial 146 points and achieving a record 15 wins in a single season. Notably, Ferrari improved their standing with Leclerc finishing second, and Red Bull maintained their position as the team to beat with Perez also featuring in the top three. The 2022 season's F1 regulations, which introduced major changes to the car design with an emphasis on ground-effect aerodynamics, appear to have shaken up the order slightly. This aimed to allow cars to follow each other more closely and to increase overtaking opportunities. The impact of these changes can be inferred from the reshuffling seen in team performance. Mercedes, for example, slipped from the top spot as Ferrari emerged as a stronger challenger to Red Bull. The midfield battle also saw shifts, with drivers like Ocon and Alonso gaining points, indicating that the new regulations might have provided an opportunity for teams to reset and challenge the established order.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-standings2022
#| tbl-cap: F1 Driver Standings 2022
library(tidyverse)
library(readr)
library(dplyr)
library(kableExtra)

standings_2022 <- read_csv("../data/analysis_data/standings_2022_cleaned.csv")

standings_2022 |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-standings2021
#| tbl-cap: F1 Driver Standings 2021
library(tidyverse)
library(readr)
library(dplyr)
library(kableExtra)

standings_2021 <- read_csv("../data/analysis_data/standings_2021_cleaned.csv")

standings_2021 |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

\newpage

## Method of Analysis

To analyze the impact of the F1 Regulations, we are going to be comparing the lap times and speeds at the Saudi Arabia Jeddah Corniche Circuit (a top-speed track) and the Netherlands' Circuit Zandvoort (a high-downforce track) as seen in @tbl-circuits for the 2021 and 2022 seasons, a structured approach will be employed. Initially, data will be gathered using the f1dataR package, which extracts detailed information about lap times and speeds from the Ergast API. This dataset will include drivers' performances during both qualifying and race sessions.

The analysis will be conducted between the two tracks and between the two years. This involves comparing average speeds and lap times to determine how the distinct characteristics of each track influence performance and how performance metrics have evolved from one year to the next. To enhance the method of analysis for comparing lap times and speeds at the Jeddah Corniche Circuit and Circuit Zandvoort over the 2021 and 2022 Formula 1 seasons, we will focus on three top drivers: Max Verstappen, Lewis Hamilton, and Charles Leclerc. Including these specific drivers in the study adds a layer of depth by enabling comparisons not just between tracks and years, but also among leading competitors known for their distinctive driving styles and performance under varying circuit conditions.

The analysis will proceed by segmenting the data to include only the laps completed by Verstappen, Hamilton, and Leclerc. This will allow us to conduct a detailed examination of how each driver's performance varies across the two contrasting track types and over the two seasons.

Finally, the results will be visually represented through various graphs and plots to illustrate variability and distribution. This focused approach on Verstappen, Hamilton, and Leclerc will provide richer insights into driver-specific strategies and skills, adding substantial value to the overall analysis of Formula 1 performances across varied circuit designs. This comparative layer will help in understanding how individual capabilities and team strategies play out in high-speed vs. high-downforce conditions, thereby enriching the conclusions drawn from the data.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-circuits
#| tbl-cap: F1 Circuits Analysis
library(tidyverse)
library(kableExtra)

variable_data <- data.frame( 
  CircuitName = c("Saudi Arabia Jeddah Corniche Circuit", 
               "Netherlands' Circuit Zandvoort"), 
  TrackType = c("Top Speed",
                  "High Downforce"), 
  TyreWear = c("High", 
              "Low")
)
variable_data |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

## Data Limitations

While the f1dataR [@f1dataR] package offers valuable resources for analyzing Formula 1 data, it does come with certain limitations that users need to consider. First, the quality and completeness of the data are dependent on the Ergast API, which means any missing or inaccurately recorded data in the API will be reflected in the outputs obtained through f1dataR. Additionally, the API's update frequency might not align with real-time event conclusions, potentially leading to delays in accessing the most current race data. The historical data provided may also have gaps, particularly from the earlier years of Formula 1 racing, which could limit analyses of long-term trends or comparisons over extended periods. Moreover, the package may not include certain types of data, such as detailed telemetry or in-depth weather conditions, which could be crucial for more nuanced analyses or predictive modeling. Therefore, while f1dataR facilitates access to a wealth of F1 data, users must be mindful of these constraints when conducting comprehensive or real-time analyses.

## Data Cleaning

R [@citeR] was the language and environment used for this paper as well as throughout the data cleaning process, with different packages such as tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], tidyr [@tidyr], knitr [@knitr], janitor [@janitor], lubridate [@lubridate] packages.

The data cleaning process for these datasets involves several steps to ensure the data is accurate and suitable for analysis. Initially, the dataset's structure is reviewed to verify data types and identify any inconsistencies. This is crucial for understanding the data's composition and guiding subsequent transformations. Key transformations include converting relevant columns to appropriate data types, such as changing lap times to numeric values for analytical operations and categorizing seasons as factors for grouped analyses. Such transformations are essential for maintaining data integrity and ensuring computational accuracy. A critical cleaning step involves filtering out anomalies, specifically laps exceeding 100 seconds, based on the assumption that these represent non-standard race conditions like pit stops or errors. This step helps focus the dataset on typical racing performance. Further, the dataset undergoes a purging of incomplete records, removing any rows with missing data to enhance the robustness of the statistical analysis. Finally, renaming columns to more descriptive terms facilitates easier understanding and manipulation of the data, aligning with best practices in data documentation and usability.

# Results {#sec-results}

## Lap Time Comparison Between 2021 and 2022 in Jeddah - Max Verstappen

@fig-maxjeddah presents Max Verstappen's lap times in Jeddah for two consecutive Formula 1 seasons, 2021 and 2022. The laps are numbered from 1 to over 50 on the horizontal axis, while the vertical axis shows the lap time in seconds. In 2021, denoted by the blue line, there's a sharp decrease in lap times from the first lap, stabilizing around the 93-second mark, suggesting a rapid adjustment to the track or conditions before finding consistency. Notably, there's less variation in lap times for this season, with most laps being clustered closely around the 93 to 95-second range, indicating a relatively steady performance.

In contrast, the 2022 season, shown in red, exhibits more variability, with lap times generally higher and more erratic. Starting around the 97-second mark, these lap times show greater fluctuations, including some particularly slow laps that spike above 95 seconds, which could indicate instances of traffic, mistakes, or other delays. Overall, the performance in 2022 seems to be less consistent and slower compared to 2021. The distinct difference in patterns between the two seasons could be attributed to changes in car performance, track conditions, strategic decisions, or regulations affecting the car's capabilities.

```{r}
#| label: fig-maxjeddah
#| fig-cap: Max Verstappen Lap Times in Jeddah - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ver_jeddah_2022 <- read_csv("../data/analysis_data/verstappen_laps_jeddah_2022_cleaned.csv")
ver_jeddah_2021 <- read_csv("../data/analysis_data/verstappen_laps_jeddah_2021_cleaned.csv")

# Add a season column to each dataframe
ver_jeddah_2021$Season <- '2021'
ver_jeddah_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ver_jeddah_2021, ver_jeddah_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "blue", "2022" = "red"))

```

\newpage

## Lap Time Comparison Between 2021 and 2022 in Jeddah - Lewis Hamilton

@fig-lewisjeddah tracks Lewis Hamilton's lap times at the Jeddah circuit over two different seasons, 2021 and 2022. In 2021, represented by the purple line, we observe a swift decrease in lap times at the start, steadying to just below 93 seconds and maintaining a relatively consistent pace with minor fluctuations throughout the race. The 2021 data suggests a quick adaptation to track conditions, with Hamilton managing to keep a tight window of performance throughout his laps.

Comparatively, the 2022 season, indicated by the yellow line, shows a different performance profile. The lap times start higher than in 2021, showing Hamilton's laps initially hover around 97 seconds. There is a notable increase in variability, with spikes suggesting laps significantly slower than the general trend. These spikes could imply various events, such as dealing with traffic, pitting, or possible errors. Towards the latter part of the graph, the lap times improve, hinting at a potential adaptation or strategic changes during the race.

The broader range of lap times in 2022 and the overall slower laps compared to 2021 might be influenced by the aforementioned regulation changes. These changes could have affected car handling and performance characteristics, necessitating a period of adaptation. Additionally, it’s possible that the 2022 Mercedes car may not have been as well-suited to the demands of the Jeddah circuit as the previous year's model, or that team strategies were different in response to the new regulations and competition.

```{r}
#| label: fig-lewisjeddah
#| fig-cap: Lewis Hamilton Lap Times in Jeddah - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ham_jeddah_2022 <- read_csv("../data/analysis_data/hamilton_laps_jeddah_2022_cleaned.csv")
ham_jeddah_2021 <- read_csv("../data/analysis_data/hamilton_laps_jeddah_2021_cleaned.csv")

# Add a season column to each dataframe
ham_jeddah_2021$Season <- '2021'
ham_jeddah_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ham_jeddah_2021, ham_jeddah_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))
```

\newpage

## Lap Time Comparison Between 2021 and 2022 in Jeddah - Charles Leclerc

@fig-lecjeddah illustrates Charles Leclerc's lap times at the Jeddah circuit, contrasting his performance across the 2021 and 2022 Formula 1 seasons. In the 2021 season, shown in red, we observe an initial spike in lap times, which quickly reduces and then follows a trend with some fluctuation around the 95-second mark, with several outliers suggesting instances of slower laps. The spread of lap times in 2021 is relatively broad, which may reflect varying conditions or incidents during the race that affected performance.

For the 2022 season, represented by the green line, the initial lap times are also high but quickly improve. After the initial laps, the times are mostly clustered around the 95-second mark, similar to the 2021 season, but with less variability. Notably, towards the end of the session, there is a sharp drop in lap times followed by a significant spike. The less volatile lap times throughout most of the race compared to 2021 might suggest better consistency or adaptation to the 2022 car's handling.

It's evident that in both seasons, there are moments of performance drops, which are likely due to non-competitive laps such as pit stops or laps following safety car periods. The general improvement in consistency in 2022, despite the new regulations, could indicate that Leclerc and his team were able to adapt effectively to the changes, possibly benefiting from the new aerodynamics rules designed to aid in car stability and handling. The sharp variations, especially in the 2022 data, could be indicative of specific incidents affecting those laps or strategic choices that temporarily influenced lap time performance.

```{r}
#| label: fig-lecjeddah
#| fig-cap: Charles Leclerc Lap Times in Jeddah - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

lec_jeddah_2022 <- read_csv("../data/analysis_data/leclerc_laps_jeddah_2022_cleaned.csv")
lec_jeddah_2021 <- read_csv("../data/analysis_data/leclerc_laps_jeddah_2021_cleaned.csv")

# Add a season column to each dataframe
lec_jeddah_2021$Season <- '2021'
lec_jeddah_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(lec_jeddah_2021, lec_jeddah_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))
```

\newpage

## Lap Time Comparison Between 2021 and 2022 in the Netherlands - Max Verstappen

@fig-maxdutch compares lap times achieved by Max Verstappen in the Netherlands for two consecutive Formula 1 seasons, 2021 and 2022. From the pattern of the data points, it can be observed that the majority of lap times for both years are closely clustered at the lower end of the lap time scale, which indicates consistent performance across the majority of the laps.

In comparing the two seasons, the 2022 lap times appear to be more volatile with higher spikes, suggesting that there were more instances where Verstappen had to slow down significantly. This might be reflective of the effects of the new regulations on race strategy and car performance. Additionally, the 2022 data points generally seem to trend slightly higher than those of 2021, which imply that the lap times were slower in 2022 on average.

```{r}
#| label: fig-maxdutch
#| fig-cap: Max Verstappen Lap Times in the Netherlands - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ver_dutch_2022 <- read_csv("../data/analysis_data/verstappen_laps_dutch_2022_cleaned.csv")
ver_dutch_2021 <- read_csv("../data/analysis_data/verstappen_laps_dutch_2021_cleaned.csv")

# Add a season column to each dataframe
ver_dutch_2021$Season <- '2021'
ver_dutch_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ver_dutch_2021, ver_dutch_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "blue", "2022" = "red"))

```

## Lap Time Comparison Between 2021 and 2022 in the Netherlands - Lewis Hamilton

@fig-lewisdutch shows Lewis Hamilton's lap times in the Netherlands for the 2021 and 2022 Formula 1 seasons. Looking at the 2021 season, represented by the purple dots, we see a consistent baseline of lap times with occasional spikes. The baseline lap time is quite stable, suggesting consistent performance throughout the race when unaffected by external factors. The 2022 season, shown in yellow, follows a similar pattern of consistency but with fewer and less pronounced spikes. Moreover, the baseline lap times in 2022 appear to be consistently higher than in 2021. This implies that laptimes were also slower in 2022.

```{r}
#| label: fig-lewisdutch
#| fig-cap: Lewis Hamilton Lap Times in the Netherlands - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

ham_dutch_2022 <- read_csv("../data/analysis_data/hamilton_laps_dutch_2022_cleaned.csv")
ham_dutch_2021 <- read_csv("../data/analysis_data/hamilton_laps_dutch_2021_cleaned.csv")

# Add a season column to each dataframe
ham_dutch_2021$Season <- '2021'
ham_dutch_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(ham_dutch_2021, ham_dutch_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))
```

## Lap Time Comparison Between 2021 and 2022 in the Netherlands - Charles Leclerc

@fig-lecdutch shows Charles Leclerc's lap times in the Netherlands for the 2021 and 2022 Formula 1 seasons. Each point represents the lap time for a specific lap in the race, with 2021 lap times in red and 2022 in green.

The data for both seasons exhibits a relatively stable baseline with periodic spikes. The consistency of the baseline indicates Leclerc’s general pace under normal racing conditions, while the spikes represent laps where his pace was significantly slower, likely due to pit stops, on-track incidents, or caution periods.

Comparing the two seasons, the 2022 green dots display a trend of marginally lower lap times than in 2021, suggesting an improved performance in the 2022 season. It's noteworthy that in both seasons, the spike patterns are somewhat similar, indicating that the race circumstances that caused these laps to be slower were likely regular race events like pit stops. The spikes are also comparable in magnitude between the two years, which suggests that the time spent on these slower laps did not change dramatically from year to year.

```{r}
#| label: fig-lecdutch
#| fig-cap: Charles Leclerc Lap Times in the Netherlands - 2021 vs. 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)

lec_dutch_2022 <- read_csv("../data/analysis_data/leclerc_laps_dutch_2022_cleaned.csv")
lec_dutch_2021 <- read_csv("../data/analysis_data/leclerc_laps_dutch_2021_cleaned.csv")

# Add a season column to each dataframe
lec_dutch_2021$Season <- '2021'
lec_dutch_2022$Season <- '2022'

# Combine the two dataframes
combined_df <- rbind(lec_dutch_2021, lec_dutch_2022)

# Create the plot
ggplot(combined_df, aes(x = `Lap #`, y = `LapTime in Seconds`, color = Season)) +
  geom_line() + # This will connect points with lines, showing the trend
  geom_point() + # This will plot individual lap times as points
  theme_minimal() + # A clean, minimal theme for the plot
  labs(
       x = "Lap Number",
       y = "Lap Time (Seconds)",
       color = "Season") +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))
```

\newpage

## Fastest Qualifying Lap Comparison Between 2021 and 2022 in Jeddah - Max Verstappen, Lewis Hamilton and Charles Leclerc

@fig-verspeed-jeddah, @fig-hamspeed-jeddah, @fig-lecspeed-jeddah show Max Verstappen, Lewis Hamilton and Charles Leclerc's speed comparison during qualifying sessions in Jeddah for the 2021 and 2022 Formula 1 seasons.

The top graph shows a general speed comparison throughout a single lap. Both lines (2021, 2022) display similar shapes, which means that the speed profiles for the lap are fairly consistent from year to year, suggesting that the general flow and demands of the track haven't changed dramatically. However, the 2022 line largely sits above the 2021 line, indicating that the drivers speeds were generally higher during the qualifying lap in the 2022 season.

The bottom graph zeroes in on the speeds at the final corner of the lap which is a very high downforce corner. It's noticeable that the 2022 speed starts higher but then drops below the 2021 speed as the corner progresses. This suggests in terms of straight-line speed, the 2022 cars benefit from a reduction in drag due to sleeker aerodynamic profiles and simpler wing designs. This reduction allows the cars to achieve higher top speeds, as drag is a key limiting factor on straights. The graph reflecting the drivers speed during a qualifying lap in Jeddah corroborates this, with the 2022 car showing a generally higher velocity compared to the 2021 car, suggesting a possible increase in aerodynamic efficiency.

However, this enhanced straight-line speed contrasts with the performance in corners. The heavier 2022 cars, combined with the sensitivity of ground effect aerodynamics to ride height changes—which are more pronounced at lower speeds—result in a compromise in agility and cornering speed. The detailed segment of the drivers lap times at the final corner in Jeddah illustrates this, where the 2022 car initially enters the corner at a higher speed, indicative of the benefits on the preceding straight, but then experiences a sharper decline in speed through the corner itself. This suggests that while the car gains time on the straights, it loses some of this advantage in the slower sections of the track due to the new aerodynamic characteristics and increased weight.

```{r}
#| label: fig-verspeed-jeddah
#| fig-cap: Verstappen Speed Comparison - Qualifying Jeddah 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

ver_telemetry_data_jeddah_2022 <- read.csv("../data/raw_data/verstappen_telemetry_jeddah.csv")
ver_telemetry_data_jeddah_2021 <- read.csv("../data/raw_data/verstappen_telemetry_jeddah_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
ver_telemetry_data_jeddah_2022$season <- "2022"
ver_telemetry_data_jeddah_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(ver_telemetry_data_jeddah_2022, ver_telemetry_data_jeddah_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
plot1 <- ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "blue"))

# Filter data around 60 seconds (between 55 and 65 seconds)
filtered_data <- combined_data %>%
  filter(time >= 55 & time <= 65)

# Plotting
plot2 <- ggplot(filtered_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022 at Final Corner",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "blue"))

grid.arrange(plot1, plot2, ncol = 1, nrow = 2)

```

```{r}
#| label: fig-hamspeed-jeddah
#| fig-cap: Hamilton Speed Comparison - Qualifying Jeddah 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

ham_telemetry_data_jeddah_2022 <- read.csv("../data/raw_data/hamilton_telemetry_jeddah.csv")
ham_telemetry_data_jeddah_2021 <- read.csv("../data/raw_data/hamilton_telemetry_jeddah_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
ham_telemetry_data_jeddah_2022$season <- "2022"
ham_telemetry_data_jeddah_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(ham_telemetry_data_jeddah_2022, ham_telemetry_data_jeddah_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
plot1 <- ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))

# Filter data around 60 seconds (between 55 and 65 seconds)
filtered_data <- combined_data %>%
  filter(time >= 55 & time <= 65)

# Plotting
plot2 <- ggplot(filtered_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022 at Final Corner",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))

grid.arrange(plot1, plot2, ncol = 1, nrow = 2)

```

```{r}
#| label: fig-lecspeed-jeddah
#| fig-cap: Leclerc Speed Comparison - Qualifying Jeddah 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

lec_telemetry_data_jeddah_2022 <- read.csv("../data/raw_data/leclerc_telemetry_jeddah.csv")
lec_telemetry_data_jeddah_2021 <- read.csv("../data/raw_data/leclerc_telemetry_jeddah_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
lec_telemetry_data_jeddah_2022$season <- "2022"
lec_telemetry_data_jeddah_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(lec_telemetry_data_jeddah_2022, lec_telemetry_data_jeddah_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
plot1 <- ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))

# Filter data around 60 seconds (between 55 and 65 seconds)
filtered_data <- combined_data %>%
  filter(time >= 55 & time <= 65)

# Plotting
plot2 <- ggplot(filtered_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022 at Final Corner",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))

grid.arrange(plot1, plot2, ncol = 1, nrow = 2)
```

\newpage

## Fastest Qualifying Lap Comparison Between 2021 and 2022 in the Netherlands - Max Verstappen, Lewis Hamilton and Charles Leclerc

@fig-verspeed-netherlands, @fig-hamspeed-netherlands, @fig-lecspeed-netherlands illustrate a comparison of Max Verstappen, Lewis Hamilton and Charles Leclerc's speeds during qualifying sessions in the Netherlands for the 2021 and 2022 Formula 1 seasons, giving insight into the performance changes due to new regulations.

Like the previous graphs, the top graph shows a general comparison of speeds throughout a lap, marked by time in seconds. The lines trace the speed of the car at different points of the lap, with two different lines, one representing 2021 and the other for 2022. We observe that the speed profiles for both years are quite similar in their fluctuations, which correspond to the acceleration and deceleration points around the lap. However, the line for 2022 occasionally edges above the line from 2021, especially in the second half of the graph. This indicates that driver's car in 2022 was able to achieve higher speeds at certain points during the lap, potentially benefiting from the reduced aerodynamic drag on the straights due to the new regulations.

The bottom graph provides a focused comparison of speeds at a critical section of the track involving turns 11 and 12. The 2022 speeds dip below those of 2021 as the car approaches and navigates through the corners. This supports the notion that while the 2022 F1 cars may gain speed on the straights, they are likely to lose some of that advantage in slower, high-downforce corners. The increased weight of the 2022 cars, coupled with the change to a ground-effect aerodynamic model that produces less downforce at low speeds, could result in slower cornering speeds. This slower speed through corners can be critical in sectors of the track where maintaining momentum is key, reflecting the trade-offs introduced by the new regulations.

Overall, these graphs suggest that the 2022 F1 cars could exhibit a complex performance profile that is a direct consequence of the trade-offs between straight-line speed and cornering agility introduced by the regulatory changes. The driver's telemetry data provides a practical example of these effects, showing possible gains on the straights and challenges in the corners.

```{r}
#| label: fig-verspeed-netherlands
#| fig-cap: Verstappen Speed Comparison - Qualifying Netherlands 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

ver_telemetry_data_dutch_2022 <- read.csv("../data/raw_data/verstappen_telemetry_dutch.csv")
ver_telemetry_data_dutch_2021 <- read.csv("../data/raw_data/verstappen_telemetry_dutch_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
ver_telemetry_data_dutch_2022$season <- "2022"
ver_telemetry_data_dutch_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(ver_telemetry_data_dutch_2022, ver_telemetry_data_dutch_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
plot1 <- ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "blue"))

filtered_data <- combined_data %>%
  filter(time >= 45 & time <= 60)

# Plotting
plot2 <- ggplot(filtered_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022 at Turn 11 and 12",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "blue"))

grid.arrange(plot1, plot2, ncol = 1, nrow = 2)
```

```{r}
#| label: fig-hamspeed-netherlands
#| fig-cap: Hamilton Speed Comparison - Qualifying Netherlands 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

ham_telemetry_data_dutch_2022 <- read.csv("../data/raw_data/hamilton_telemetry_dutch.csv")
ham_telemetry_data_dutch_2021 <- read.csv("../data/raw_data/hamilton_telemetry_dutch_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
ham_telemetry_data_dutch_2022$season <- "2022"
ham_telemetry_data_dutch_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(ham_telemetry_data_dutch_2022, ham_telemetry_data_dutch_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
plot1 <- ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))

filtered_data <- combined_data %>%
  filter(time >= 45 & time <= 60)

# Plotting
plot2 <- ggplot(filtered_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022 at Turn 11 and 12",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "purple", "2022" = "gold"))

grid.arrange(plot1, plot2, ncol = 1, nrow = 2)
```

```{r}
#| label: fig-lecspeed-netherlands
#| fig-cap: Leclerc Speed Comparison - Qualifying Netherlands 2022 vs 2021
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

lec_telemetry_data_dutch_2022 <- read.csv("../data/raw_data/leclerc_telemetry_dutch.csv")
lec_telemetry_data_dutch_2021 <- read.csv("../data/raw_data/leclerc_telemetry_dutch_2021.csv")

# Assuming telemetry_data_2022 and telemetry_data_2021 are your datasets
lec_telemetry_data_dutch_2022$season <- "2022"
lec_telemetry_data_dutch_2021$season <- "2021"

# Combine the datasets
combined_data <- rbind(lec_telemetry_data_dutch_2022, lec_telemetry_data_dutch_2021)

# Convert session_time to a numeric type if not already
combined_data$session_time <- as.numeric(combined_data$session_time)

# Plotting
plot1 <- ggplot(combined_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))

filtered_data <- combined_data %>%
  filter(time >= 45 & time <= 60)

# Plotting
plot2 <- ggplot(filtered_data, aes(x = time, y = speed, color = season)) +
  geom_line() +
  labs(title = "Comparison of Speeds: 2021 vs 2022 at Turn 11 and 12",
       x = "Session Time (s)",
       y = "Speed (km/h)") +
  theme_minimal() +
  scale_color_manual(values = c("2021" = "red", "2022" = "darkgreen"))

grid.arrange(plot1, plot2, ncol = 1, nrow = 2)
```

\newpage

## Comparison of Number of Overtakes throughout each Season 2021 vs 2022

@fig-overtakes compares the number of overtakes during the Formula 1 racing seasons of 2021 and 2022, with Chart A focusing on total overtakes and Chart B on overtakes shown on TV. In both charts, there is a noticeable increase in the number of overtakes in the 2022 season compared to 2021.

```{r}
#| label: fig-overtakes
#| fig-cap: Number of Overtakes Throughout Each Season 2021 v 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

data <- read.csv("../data/raw_data/Overtaking_Season_Averages.csv")

# Filter data for the years 2021 and 2022
filtered_data <- data[data$Season %in% c(2021, 2022), ]

plot1 <- ggplot(filtered_data, aes(x = as.factor(Season), y = Overtakes, fill = "Total Overtakes")) +
    geom_bar(stat = "identity", color = "black", fill = "blue") +
    labs(
         x = "Season", y = "Number of Overtakes") +
    theme_minimal()+
  labs(tag = "A")

plot2 <- ggplot(filtered_data, aes(x = as.factor(Season), y = Overtakes.on.TV, fill = "Overtakes on TV")) +
    geom_bar(stat = "identity", color = "black", fill = "red") +
    labs(
         x = "Season", y = "Number of Overtakes on TV") +
    theme_minimal()+
  labs(tag = "B")

grid.arrange(plot1, plot2, ncol = 2, nrow = 1)

```

\newpage

## Comparison of Average Number of Overtakes throughout each Season 2021 vs 2022

@fig-averageovertakes represent the average number of overtakes per race in the Formula 1 seasons of 2021 and 2022. Chart A shows a slight increase in the average number of overtakes from the 2021 to the 2022 season, while Chart B illustrates a more noticeable increase in the average number of overtakes broadcast on TV during the same period.

```{r}
#| label: fig-averageovertakes
#| fig-cap: Average Number of Overtakes Per Race 2021 v 2022
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)

data <- read.csv("../data/raw_data/Overtaking_Season_Averages.csv")

# Filter data for the years 2021 and 2022
filtered_data <- data[data$Season %in% c(2021, 2022), ]

plot1 <- ggplot(filtered_data, aes(x = as.factor(Season), y = Average, fill = "Average")) +
    geom_bar(stat = "identity", color = "black", fill = "purple") +
    labs(
         x = "Season", y = "Average Number of Overtakes") +
    theme_minimal()+
  labs(tag = "A")

plot2 <- ggplot(filtered_data, aes(x = as.factor(Season), y = Average.on.TV, fill = "Average on TV")) +
    geom_bar(stat = "identity", color = "black", fill = "green") +
    labs(
         x = "Season", y = "Average Number of Overtakes on TV") +
    theme_minimal()+
  labs(tag = "B")

grid.arrange(plot1, plot2, ncol = 2, nrow = 1)

```

## How the regulations had an effect on total and average number of overtakes

This increase can be attributed to the major regulatory changes introduced in 2022, which aimed at promoting closer racing and more overtaking opportunities. The 2022 regulations involved a shift to ground-effect aerodynamics, allowing cars to follow each other more closely without losing significant downforce, which is crucial for maintaining speed in corners. Additionally, changes were made to the car's wings and wheel design to reduce the aerodynamic wake that can disturb the airflow for following cars, further facilitating overtaking maneuvers.

As a result, the data suggests that these regulations had their intended effect, leading to more competitive races with increased overtaking, which is reflected in both the total number of overtakes and those broadcast on television. The changes not only improved the chances of overtaking but also likely enhanced the viewing experience for fans, as reflected in the consistency of overtakes shown on TV, which closely aligns with the actual increase in overtakes. This might indicate that more of the action was deemed broadcast-worthy, potentially due to the more dynamic and visually engaging nature of the overtakes under the new regulations.

# Discussion {#sec-discussion}

## Are the 2022 f1 cars faster then their predecessor?

The 2022 regulations introduced a shift back to ground effect aerodynamics, a method that relies on the car’s underbody design to generate a large portion of its downforce. This allows for more stable aerodynamics when following another car closely, which ideally enhances overtaking opportunities and racing quality. However, these changes also mean that the cars generate less downforce overall compared to the high-rake aerodynamics used in previous years [@f1_changes]. Less downforce generally leads to slower speeds in high-speed corners, which can contribute to slower lap times overall which is why we see that Max Verstappen, Lewis Hamilton and Charles Leclerc all had relatively slower laptimes from 2021 to 2022.

Another significant factor is the increase in the minimum weight of the cars. The 2022 regulations raised the minimum weight from 748 kg to 798 kg [@f1_weight], largely due to the introduction of heavier and safer chassis requirements and other components like bigger wheels (18 inches up from 13 inches) [@f1_changes]. Heavier cars require more energy to accelerate and decelerate, which naturally results in slower lap times, especially evident in tighter, more technical circuits where agility and quick changes of direction are crucial.

The switch to 18-inch wheels from the previous 13-inch also impacts the car's performance. The larger rims come with a lower-profile tire, which behaves differently compared to the high-profile tires used earlier. These changes affect the suspension geometry and dynamics, as well as how the tire copes with heat and wear during a race [@f1_changes]. While designed to be more durable, the new tires may not provide the same level of grip and responsiveness as the previous designs, potentially leading to slower lap times. 2022 F1 cars are engineered with a different set of priorities and constraints, focusing on enhancing the racing experience and safety [@f1_changes]. While this might mean they are initially slower in terms of lap times, they offer improved performance in terms of race dynamics, such as in duels and overtakes.

## Deeper Insight into the 2022 Regulation Changes

The 2022 Formula 1 regulations demonstrate how precise, targeted regulatory changes can substantially enhance the dynamics of competition within sports. These regulations primarily focused on redesigning the aerodynamics of the vehicles, aiming to mitigate the 'dirty air' effect—turbulent air left by a leading car that significantly reduces the aerodynamic efficiency of following cars. By introducing cars with ground effect designs that rely less on aerodynamic appendages susceptible to turbulence, the regulations reduced the aerodynamic wake and thereby allowed trailing cars to follow more closely without significant loss of performance.

The impact of these changes on the sport is observable through a marked increase in the number and quality of overtaking maneuvers across various circuits, especially in high-speed corners where following closely used to be particularly challenging. @fig-averageovertakes and @fig-overtakes shows not only a quantitative increase in overtakes but also qualitative improvements in race tactics, as drivers now exploit new overtaking opportunities that previously did not exist. This has led to races that are less predictable and more dynamic, contributing to a more engaging viewer experience and broader fan engagement.

Moreover, these changes underscore a deeper principle that well-conceived regulations can drive innovation and elevate performance standards. In Formula 1, teams have been pushed to innovate within the new aerodynamic constraints, leading to rapid advancements in related technologies. This progress not only enhances the sport but also trickles down to automotive engineering more broadly, potentially influencing road car designs and technologies. Thus, the F1 2022 regulations illustrate how strategic regulatory frameworks can simultaneously enhance competitive balance, improve spectator value, and spur broader technological innovation within an industry.

## Environmental and Economic Impact of 2022 Regulation Changes

The 2022 Formula 1 regulations bring into focus the economic and environmental dimensions of sports regulation, particularly in how they adapt to contemporary sustainability challenges. The introduction of a cost cap is a pivotal development as it aims to equalize the competitive environment by limiting the amount teams can spend throughout a season [@f1_costcap]. This economic mechanism is designed not only to stop the escalating expenditures that often see wealthier teams pull far ahead but also to create a more strategically diverse field where innovation must be achieved within fixed financial boundaries. This could lead to smarter, more efficient use of resources, which is essential for the sport's long-term sustainability.

On the environmental front, the regulations mandate the use of E10 fuel, integrating a 10% ethanol blend derived from renewable sources [@f1_fuel]. This move aligns with global environmental goals by reducing the reliance on pure fossil fuels and cutting down on the sport's overall carbon emissions. Furthermore, this shift is accompanied by more rigorous aerodynamic regulations intended to decrease downforce and thus reduce fuel consumption during races. These measures collectively demonstrate Formula 1’s commitment to playing a proactive role in addressing climate change, offering a model for other sectors where environmental impact and high-performance demands intersect [@f1_fuel].

These changes not only reflect an awareness of Formula 1's environmental footprint but also its role in pioneering technologies that can have broader applications. As such, the sport's regulatory framework serves as a test bed for sustainable practices that could influence automotive design and fuel technology in commercial industries, underscoring the potential of sports as a catalyst for environmental innovation.

## Potential Issues with New Regulations

The 2022 Formula 1 regulations, while innovative and well-intentioned in their approach to improving race competitiveness and sustainability, manifest certain weaknesses that could undermine their effectiveness. A primary concern is the reliance on relatively short-term data to assess the impact of these regulations. This limited timeframe may not adequately reflect the full range of effects, particularly in how teams adapt strategically over longer periods. Such an evaluation might overlook emerging trends or delayed consequences, leading to potentially skewed interpretations of the regulations' success or failure.

Another notable issue is the increased complexity and cost associated with the new technical requirements, despite the implementation of cost caps [@f1_costcap]. These regulations demand significant changes to car designs, particularly in aerodynamics and tire usage, which can disproportionately affect smaller teams with fewer resources for rapid adaptation and innovation. This situation can paradoxically widen the competitive gap the regulations aim to narrow.

Moreover, while the switch to ground effect aerodynamics aims to reduce the 'dirty air' problem, it simultaneously introduces new challenges in handling and tire degradation, potentially leading to unanticipated competitive disparities and affecting the overall quality of racing.

## How should F1 Proceed in the Future

As Formula 1 continues to evolve under the 2022 regulations, several areas require further exploration to fully understand and optimize the impacts of these changes. First and foremost, a longer-term assessment is crucial. The initial seasons under new regulations provide valuable data, but the true test will come as teams fully adapt and mature within the new framework. This will allow for a more accurate evaluation of the regulations' long-term sustainability and their effect on the competitive balance. Investigating over multiple seasons will help determine if smaller teams can indeed close the gap with top-tier teams or if financial and technical disparities continue to dominate outcomes.

Additionally, further research into the environmental impacts of the sport is needed, particularly concerning the real-world benefits of using sustainable fuels and advanced materials. As the global emphasis on reducing carbon footprints intensifies, F1 could serve as a pioneering platform for innovative green technologies. This includes not only the fuels used but also the materials and methods employed in manufacturing and transporting race equipment.

Technological development should also be a focal point. The potential of emerging technologies such as artificial intelligence and advanced simulation tools to enhance car performance and race strategies should be explored. These technologies could help teams optimize their designs more efficiently and adapt to regulations more effectively.

Furthermore, enhancing fan engagement through adjustments in race formats, accessibility of content, and interactive platforms will be vital. As the sport aims to attract a broader, more diverse audience, understanding the preferences and consumption patterns of this audience will be essential.

# Conclusion

The 2022 Formula 1 regulations represent a pivotal moment in the sport’s history, aiming to enhance competitiveness and sustainability while addressing both the environmental and economic challenges of modern motorsport. These regulations have introduced significant changes in car design, notably through ground effect aerodynamics, and have implemented a cost cap that levels the playing field, allowing for a more diverse and competitive environment. Preliminary observations indicate that these changes have indeed increased the quality of racing, as evidenced by more frequent and strategic overtaking maneuvers and closer competition among teams.

My analysis began with that the 2022 cars would exhibit faster straight-line speeds due to reduced drag but suffer in cornering speeds due to less downforce. The data collected and analyzed, confirmed this. All drivers showed an increase in top speed from 2021 to 2022, underscoring the effects of aerodynamic simplification. Yet, regardless of the type of corner whether high-speed or low-speed, all tracks examined demonstrated an increase in lap times. This suggests that while the cars gained speed in the straights, they lost substantial time in the corners, leading to overall slower lap times across a full circuit.

Despite these findings, the regulations have positively influenced the sport by enhancing race competitiveness and overtaking, as well as by advancing environmental considerations through the use of sustainable fuels and more stringent aerodynamic rules. As Formula 1 continues to innovate within these new guidelines, it maintains its status as a pinnacle of motorsport technology and strategy.

Looking ahead, further studies are warranted to explore the long-term impacts of these regulations, particularly as the sport approaches another significant regulatory shift in 2026 with new engines, aerodynamics, and fully sustainable fuels. The ongoing challenge for Formula 1 will be to balance technological innovation and competitive racing with environmental sustainability and economic viability, ensuring that the sport remains at the front of global motorsport while adapting to the demands of a changing world.

\newpage

# References